---
title: "HiSSEChecks"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

```{r echo=FALSE, include=FALSE}
setwd("/Users/nhans/DiversificationPaper/Diversification2020/")
source("HiSSEplots-cetacean.R")

# Generating colorblind friendly palette (30)
library(RColorBrewer)
n <- 30
colrs <- brewer.pal.info[brewer.pal.info$colorblind == TRUE, ]
col_vec = unlist(mapply(brewer.pal, colrs$maxcolors, rownames(colrs)))
col <- sample(col_vec, n)
#area <- rep(1,n)
#pie(area, col = col)
```

```{r}
# Boxplots the original parameters
par(mfrow=c(2,2))
boxplot(tau1.his.clone, main=expression(bold(tau[0][A])),col=col[1],xlab="Clones")
boxplot(tau2.his.clone, main=expression(bold(tau[1][A])),col=col[2],xlab="Clones")
boxplot(tau3.his.clone, main=expression(bold(tau[0][B])),col=col[3],xlab="Clones")
boxplot(tau4.his.clone, main=expression(bold(tau[1][B])),col=col[4],xlab="Clones")
mtext(expression(bold(tau)), side =3, line = -1.5, outer = TRUE,cex = 2.5)


par(mfrow=c(2,2))
boxplot(ep1.his.clone, main=expression(bold(epsilon[0][A])),col=col[5])
boxplot(ep2.his.clone, main=expression(bold(epsilon[1][A])),col=col[6])
boxplot(ep3.his.clone, main=expression(bold(epsilon[0][B])),col=col[7])
boxplot(ep4.his.clone, main=expression(bold(epsilon[1][B])),col=col[8])
mtext(expression(bold(epsilon)), side =3, line = -1.5, outer = TRUE,cex = 2.5)


# Boxplots of reparameterized parameters
par(mfrow=c(2,2))
boxplot(lambda1.his.clone, main=expression(bold(lambda[0][A])),col=col[5])
boxplot(lambda2.his.clone, main=expression(bold(lambda[1][A])),col=col[6])
boxplot(lambda3.his.clone, main=expression(bold(lambda[0][B])),col=col[7])
boxplot(lambda4.his.clone, main=expression(bold(lambda[1][B])),col=col[8])
mtext(expression(bold(lambda)), side =3, line = -1.5, outer = TRUE,cex = 2.5)

par(mfrow=c(2,2))
boxplot(mu1.his.clone, main=expression(bold(mu[0][A])),col=col[5])
boxplot(mu2.his.clone, main=expression(bold(mu[1][A])),col=col[6])
boxplot(mu3.his.clone, main=expression(bold(mu[0][B])),col=col[7])
boxplot(mu4.his.clone, main=expression(bold(mu[1][B])),col=col[8])
mtext(expression(bold(mu)), side =3, line = -1.5, outer = TRUE,cex = 2.5)
```



# (I) Parameter combinations - 
## A) Turnover rates
### 1. Checks for addition $\tau_{0A}+\tau_{0B}$ AND $\tau_{1A} + \tau_{1B}$ 

My reason behind adding turnover rates for state 0A  and 0B into state 0 and state 1A and 1B to state 1, is that this should be  similar to BiSSE if we reparameterize BiSSE as $\lambda_{0} + \mu_{0}$ and $\lambda_{1} + \mu_{1}$. 
```{r tau0_tau1, echo=FALSE, warning=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B
tau_state0 <- tau1.his.clone + tau3.his.clone
tau_state1 <- tau2.his.clone + tau4.his.clone

# for 0A+0B
#png("tau_0.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(tau_state0[,count],main=k,col =cbbPalette[1] ,xlim = c(0,1),xlab=expression(bold(tau[0][A]+tau[0][B])))
count<-count+1
}
mtext(expression(bold(tau[0][A]+tau[0][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 1A+1B
#png("tau_1.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(tau_state1[,count],main=k,col =cbbPalette[2],xlim = c(0,1),xlab=expression(bold(tau[1][A]+tau[1][B])))
count<-count+1
}
mtext(expression(bold(tau[1][A]+tau[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```

### 2. Checks for addition $\tau_{0A}+\tau_{1A}$ AND $\tau_{0B} + \tau_{1B}$ 
Here I am checking if hidden states have a signal in turnover rates  i.e. if the two rate classes A and B can also be combined. Here state A is represented by the sum $\tau_{0A}+\tau_{1A}$ and state B is represented by the sum $\tau_{0B} + \tau_{1B}$. 

```{r tauA_tauB, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
tau_stateA <- tau1.his.clone + tau2.his.clone
tau_stateB <- tau3.his.clone + tau4.his.clone

# for 0A+1A
#png("tau_A.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(tau_stateA[,count],main=k,col =cbbPalette[3],xlim = c(0,1),xlab=expression(bold(tau[0][A]+tau[1][A])))
count<-count+1
}
mtext(expression(bold(tau[0][A]+tau[1][A])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 0B+1B
#png("tau_B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(tau_stateB[,count],main=k,col =cbbPalette[4],xlim = c(0,1),xlab=expression(bold(tau[0][B]+tau[1][B])))
count<-count+1
}
mtext(expression(bold(tau[0][B]+tau[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```


## B) Speciation rates
### 1. Checks for addition $\lambda_{0A}+\lambda_{0B}$ AND $\lambda_{1A} + \lambda_{1B}$ 
My reason behind adding speciation rates for state 0A  and 0B into state 0 and state 1A and 1B to state 1, is that this should reduce to BiSSE for the following reparameterization $\lambda_{0A} + \lambda_{0B}$ and $\lambda_{1A} + \lambda_{1B}$. 
```{r lambda0_lambda1, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B
lambda_state0 <- lambda1.his.clone + lambda3.his.clone
lambda_state1 <- lambda2.his.clone + lambda4.his.clone

# for 0A+0B
#png("lambda_0.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(lambda_state0[,count],main=k,col =cbbPalette[1] ,xlim = c(0,0.8),xlab=expression(bold(lambda[0][A]+lambda[0][B])))
count<-count+1
}
mtext(expression(bold(lambda[0][A]+lambda[0][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 1A+1B
#png("lambda_1.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(lambda_state1[,count],main=k,col =cbbPalette[2],xlim = c(0,0.8),xlab=expression(bold(lambda[1][A]+lambda[1][B])))
count<-count+1
}
mtext(expression(bold(lambda[1][A]+lambda[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```


### 2. Checks for addition $\lambda_{0A}+\lambda_{1A}$ AND $\lambda_{0B} + \lambda_{1B}$ 
Here I am checking if hidden states have a signal in speciation rates  i.e. if the two rate classes A and B can also be combined. Here state A is represented by the sum $\lambda_{0A}+\lambda_{1A}$ and state B is represented by the sum $\lambda_{0B} + \lambda_{1B}$. 

```{r lambdaA_lambda1, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
lambda_stateA <- lambda1.his.clone + lambda2.his.clone
lambda_stateB <- lambda3.his.clone + lambda4.his.clone

# for 0A+1A
#png("lambda_A.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(lambda_stateA[,count],main=k,col =cbbPalette[3],xlim = c(0,0.8),xlab=expression(bold(lambda[0][A]+lambda[1][A])))
count<-count+1
}
mtext(expression(bold(lambda[0][A]+lambda[1][A])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 0B+1B
#png("lambda_B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(lambda_stateB[,count],main=k,col =cbbPalette[4],xlim = c(0,0.8),xlab=expression(bold(lambda[0][B]+lambda[1][B])))
count<-count+1
}
mtext(expression(bold(lambda[0][B]+lambda[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```

## C) Extinction fractions
### 1. Checks for addition $\epsilon_{0A}+\epsilon_{0B}$ AND $\epsilon_{1A} + \epsilon_{1B}$ 
Here we are adding extinction fractions rates for state 0A  and 0B into state 0 and state 1A and 1B to state 1, is that this should be  similar to BiSSE if we reparameterize the BiSSE model in terms of extinction fraction such as   $\epsilon_{0A} + \epsilon_{0B}$ and $\epsilon_{1A} + \epsilon_{1B}$. 
```{r epsilon0_epsilon1, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B
ep_state0 <- ep1.his.clone + ep3.his.clone
ep_state1 <- ep2.his.clone + ep4.his.clone

# for 0A+0B
#png("epsilon_0.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(ep_state0[,count],main=k,col =cbbPalette[1] ,xlim = c(0,1),xlab=expression(bold(epsilon[0][A]+epsilon[0][B])))
count<-count+1
}
mtext(expression(bold(epsilon[0][A]+epsilon[0][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 1A+1B
#png("epsilon_1.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(ep_state1[,count],main=k,col =cbbPalette[2],xlim = c(0,1),xlab=expression(bold(epsilon[1][A]+epsilon[1][B])))
count<-count+1
}
mtext(expression(bold(epsilon[1][A]+epsilon[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```


### 2. Checks for addition $\epsilon_{0A}+\epsilon_{1A}$ AND $\epsilon_{0B} + \epsilon_{1B}$ 
Checking if the hidden states have a signal when combining extinction fractions  i.e. if the two rate classes A and B can also be combined. Here state A is represented by the sum $\epsilon_{0A}+\epsilon_{1A}$ and state B is represented by the sum $\epsilon_{0B} + \epsilon_{1B}$. 

```{r epsilonA_epsilonB, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
ep_stateA <- ep1.his.clone + ep2.his.clone
ep_stateB <- ep3.his.clone + ep4.his.clone

# for 0A+1A
#png("epsilon_A.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(ep_stateA[,count],main=k,col =cbbPalette[3],xlim = c(0,1),xlab=expression(bold(epsilon[0][A]+epsilon[1][A])))
count<-count+1
}
mtext(expression(bold(epsilon[0][A]+epsilon[1][A])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 0B+1B
#png("epsilon_B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(ep_stateB[,count],main=k,col =cbbPalette[4],xlim = c(0,1),xlab=expression(bold(epsilon[0][B]+epsilon[1][B])))
count<-count+1
}
mtext(expression(bold(epsilon[0][B]+epsilon[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```

## D) Extinction rates
### 1. Checks for addition $\mu_{0A}+\mu_{0B}$ AND $\mu_{1A} + \mu_{1B}$ 

I am adding the extinction rates for state 0A  and 0B into state 0 and state 1A and 1B to state 1. By doing this the model reduces to BiSSE such that  $\mu_0 = \mu_{0A} + \mu_{0B}$ and $\mu_1= \mu_{1A} + \mu_{1B}$.
```{r mu0_mu1, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B
mu_state0 <- mu1.his.clone + mu3.his.clone
mu_state1 <- mu2.his.clone + mu4.his.clone

# for 0A+0B
#png("mu_0.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(mu_state0[,count],main=k,col =cbbPalette[1] ,xlim = c(0,0.5),xlab=expression(bold(mu[0][A]+mu[0][B])))
count<-count+1
}
mtext(expression(bold(mu[0][A]+mu[0][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 1A+1B
#png("mu_1.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(mu_state1[,count],main=k,col =cbbPalette[2],xlim = c(0,0.5),xlab=expression(bold(mu[1][A]+mu[1][B])))
count<-count+1
}
mtext(expression(bold(mu[1][A]+mu[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```


### 2. Checks for addition $\mu_{0A}+\mu_{1A}$ AND $\mu_{0B} + \mu_{1B}$ 
Checking if hidden states have a signal when combining extinction rates  i.e. if the two rate classes A and B can also be combined. Here state A is represented by the sum $\mu_{0A}+\mu_{1A}$ and state B is represented by the sum $\mu_{0B} + \mu_{1B}$. 

```{r muA_muB, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
mu_stateA <- mu1.his.clone + mu2.his.clone
mu_stateB <- mu3.his.clone + mu4.his.clone

# for 0A+1A
#png("mu_A.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(mu_stateA[,count],main=k,col =cbbPalette[3],xlim = c(0,0.5),xlab=expression(bold(mu[0][A]+mu[1][A])))
count<-count+1
}
mtext(expression(bold(mu[0][A]+mu[1][A])), side =3, line = -1.5, outer = TRUE)
#dev.off()

# for 0B+1B
#png("mu_B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(mu_stateB[,count],main=k,col =cbbPalette[4],xlim = c(0,0.5),xlab=expression(bold(mu[0][B]+mu[1][B])))
count<-count+1
}
mtext(expression(bold(mu[0][B]+mu[1][B])), side =3, line = -1.5, outer = TRUE)
#dev.off()
```

# (II) Diversification rates check "r":
A) Checking if $r_{0A}/r_{1A}$ and $r_{0B}/r_{1B}$ are convergent.
i.e. diversification rate differences per state, which checks whether state 0 has a different diversification rate than state 1


B) Checking $r_{0A}/r_{0B}$ and $r_{1A}/r_{1B}$.
i.e. diversification rate differences per hidden state i.e if hidden state for state 0 is different than what is observed (separating the main effect vs. the noise)

```{r divrates, warning=FALSE,echo=FALSE}
div.rate1<-lambda1.his.clone - mu1.his.clone
div.rate2<-lambda2.his.clone - mu2.his.clone
div.rate3<-lambda3.his.clone - mu3.his.clone
div.rate4<-lambda4.his.clone - mu4.his.clone
```



## A) $r_{0A}/r_{1A}$ and $r_0B/r_1B$. - Checking convergence

```{r r0_r1, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
div_0A_1A <- div.rate1 / div.rate2
div_0B_1B <- div.rate3 / div.rate4

# for 0A/1A
#png("r_0A_1A.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_0A_1A[,count],main=k,col =cbbPalette[5], xlab=expression(bold(r[0][A]/r[1][A])))
count<-count+1
}
mtext("Diversification 0A/1A", side =3, line = -1.25, outer = TRUE)
#dev.off()

# for 0B/1B
#png("r_0B_1B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_0B_1B[,count],main=k,col =cbbPalette[6],xlab=expression(bold(r[0][B]/r[1][B])))
count<-count+1
}
mtext("Diversification 0B/1B", side =3, line = -1.25, outer = TRUE)
#dev.off()
```

## B) $r_{0A}/r_{0B}$ annd $r_{1A}/r_{1B}$ - Checking Noise
```{r rA_rB, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
div_0A_0B <- div.rate1 / div.rate3
div_1A_1B <- div.rate2 / div.rate4

# for 0A/0B
#png("r_0A_0B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_0A_0B[,count],main=k,col =cbbPalette[7], xlab=expression(bold(r[0][A]/r[0][B])))
count<-count+1
}
mtext("Diversification 0A/0B", side =3, line = -1.25, outer = TRUE)
#dev.off()

# for 1A/1B
#png("r_1A_1B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_1A_1B[,count],main=k,col =cbbPalette[8],xlab=expression(bold(r[1][A]/r[1][B])))
count<-count+1
}
mtext("Diversification 1A/1B", side =3, line = -1.25, outer = TRUE)
#dev.off()
```
# Why we might be seeing the above results
## 1. Distribution of lambda and mu and diversification rates for each clone
These boxplots show the distribution of speciation rates $\lambda_{0A},\lambda_{1A},\lambda_{0B},\lambda_{1B}$ and extinction rates $\mu_{0A},\mu_{1A},\mu_{0B},\mu_{1B}$ calculated from HiSSE's turnover rates and extinction fractions.

Please note that for some clones the extinction rate is high leading to negative diversification rates. I added a line y=0 in diversification rates distribution plot to demonstrate that point. 

```{r}
#Format  1=0A, 2=1A, 3=0B, 4=1B.

# speciation rates distribution
#png("speciationRateDistribution.png")
par(mfrow=c(2,2))
boxplot(lambda1.his.clone, main=expression(lambda[0][A]),ylab="Rates",xlab="clones")
boxplot(lambda2.his.clone, main=expression(lambda[1][A]),ylab="Rates",xlab="clones")
boxplot(lambda3.his.clone, main=expression(lambda[0][B]),ylab="Rates",xlab="clones")
boxplot(lambda4.his.clone, main=expression(lambda[1][B]),ylab="Rates",xlab="clones")
mtext("speciation rates distribution", side =3, line = -1.5, outer = TRUE)
#dev.off()

# extinction rates distribution
#png("extinctionRateDistribution.png")
par(mfrow=c(2,2))
boxplot(mu1.his.clone, main=expression(mu[0][A]),xlab="clones",ylab="Rates")
boxplot(mu2.his.clone, main=expression(mu[1][A]),xlab="clones",ylab="Rates")
boxplot(mu3.his.clone, main=expression(mu[0][B]),xlab="clones",ylab="Rates")
boxplot(mu4.his.clone, main=expression(mu[1][B]),xlab="clones",ylab="Rates")
mtext("extinction rates distribution", side =3, line = -1.5, outer = TRUE)
#dev.off()

# diversification rates distribution
#png("diversificationRateDistribution.png")
par(mfrow=c(2,2))
boxplot(div.rate1, main=expression(r[0][A]),xlab="clones",ylab="Rates")
abline(h=0)
boxplot(div.rate2, main=expression(r[1][A]),xlab="clones",ylab="Rates")
abline(h=0)
boxplot(div.rate3, main=expression(r[0][B]),xlab="clones",ylab="Rates")
abline(h=0)
boxplot(div.rate4, main=expression(r[1][B]),xlab="clones",ylab="Rates")
abline(h=0)
mtext("diversification rates distribution", side =3, line = -1.5, outer = TRUE)
#dev.off()
```

## 2.Removing the outliers 
```{r ,echo=FALSE, warning=FALSE,include=FALSE}
# Function for finding outlier position
FindOutlierPos <- function(data) {
  lowerq = quantile(data)[2]
  upperq = quantile(data)[4]
  iqr = upperq - lowerq 
  # we identify extreme outliers
  extreme.threshold.upper = (iqr * 1.5) + upperq
  extreme.threshold.lower = lowerq - (iqr * 1.5)
  result <- which(data > extreme.threshold.upper | data < extreme.threshold.lower)
}
#--------

#pos<-apply(lambda1.his.clone, 2, FindOutlierPos)


# Function for finding outliers 
outliers <- function(x) {

  Q1 <- quantile(x)[2]
  Q3 <- quantile(x)[4]
  iqr = Q3-Q1

 upper_limit = Q3 + (iqr*1.5)
 lower_limit = Q1 - (iqr*1.5)

 x > upper_limit | x < lower_limit
}
#------

# Function for removing outlier by column ( calls outlier function above)
remove_outliers <- function(df, cols = names(df)) {
  for (col in cols) {
    df <- df[!outliers(df[[col]]),]
  }
  df
}
#--------

count<-1
for (k in k.vec) {
  #print(k)
 
  
   assign(paste("div1_","clone_",k,sep=""), div.rate1[which(outliers(div.rate1[,count])==FALSE)] )
  assign(paste("div2_","clone_",k,sep = ""),div.rate2[which(outliers(div.rate2[,count])==FALSE)] )
  assign(paste("div3_","clone_",k,sep=""), div.rate3[which(outliers(div.rate3[,count])==FALSE)] )
  assign(paste("div4_","clone_",k,sep=""), div.rate4[which(outliers(div.rate4[,count])==FALSE)] )
   print(count)
   print("length of div1")
   print(length(div.rate1[which(outliers(div.rate1[,count])==FALSE)]))
  print("length of div2")
   print(length(div.rate2[which(outliers(div.rate2[,count])==FALSE)]))
   print("length of div3")
   print(length(div.rate3[which(outliers(div.rate3[,count])==FALSE)]))
   print("length of div4")
   print(length(div.rate4[which(outliers(div.rate4[,count])==FALSE)]))
   print("-------------------")
  count<-count+1
}

x1<-cbind(div1_clone_2,div1_clone_4,div1_clone_8,div1_clone_16,div1_clone_32,div1_clone_64,div1_clone_128,div1_clone_256,div1_clone_512)
x2<-cbind(div2_clone_2,div2_clone_4,div2_clone_8,div2_clone_16,div2_clone_32,div2_clone_64,div2_clone_128,div2_clone_256,div2_clone_512)
x3<-cbind(div3_clone_2,div3_clone_4,div3_clone_8,div3_clone_16,div3_clone_32,div3_clone_64,div3_clone_128,div3_clone_256,div3_clone_512)
x4<-cbind(div4_clone_2,div4_clone_4,div4_clone_8,div4_clone_16,div4_clone_32,div4_clone_64,div4_clone_128,div4_clone_256,div4_clone_512)
colnames(x1)<-k.vec
colnames(x2)<-k.vec
colnames(x3)<-k.vec
colnames(x4)<-k.vec
```

## 3. Distribution of diversification rates after removing outliers

```{r}
# Combine the above individual rates 
#png("diversificationRateDistribution_outliersRemoved.png")
par(mfrow=c(2,2))
boxplot(x1,main=expression(r[0][A]),ylab="Rates",xlab="clones",ylim=c(0,0.5))
boxplot(x2,main=expression(r[1][A]),ylab="Rates",xlab="clones",ylim=c(0,0.5))
boxplot(x3,main=expression(r[0][B]),ylab="Rates",xlab="clones",ylim=c(0,0.5))
boxplot(x4,main=expression(r[1][B]),ylab="Rates",xlab="clones",ylim=c(0,0.5))
mtext("diversification rates distribution no outliers", side =3, line = -1.5, outer = TRUE)
#dev.off()
```
### A) $r_{0A}/r_{1A}$ and $r_0B/r_1B$. - Checking convergence -Outliers removed

```{r r0_r1_NoOut, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
div_0A_1A <- x1[1:1000,]/x2[1:1000,]
div_0B_1B <- x3[1:1000,]/x4[1:1000,]

# for 0A/1A
#png("r_0A_1A.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_0A_1A[,count],main=k,col =cbbPalette[5], xlab=expression(bold(r[0][A]/r[1][A])))
count<-count+1
}
mtext("Diversification 0A/1A - no outliers", side =3, line = -1.25, outer = TRUE)
#dev.off()

# for 0B/1B
#png("r_0B_1B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_0B_1B[,count],main=k,col =cbbPalette[6],xlab=expression(bold(r[0][B]/r[1][B])))
count<-count+1
}
mtext("Diversification 0B/1B - no outliers", side =3, line = -1.25, outer = TRUE)
#dev.off()
```

## B) $r_{0A}/r_{0B}$ annd $r_{1A}/r_{1B}$ - Checking Noise - Outliers Removed
```{r rA_rB_NoOut, warning=FALSE,echo=FALSE}
#Format  1=0A, 2=1A, 3=0B, 4=1B.
div_0A_0B <- x1[1:1000,]/x3[1:1000,]
div_1A_1B <- x2[1:1000,]/x4[1:1000,]

# for 0A/0B
#png("r_0A_0B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_0A_0B[,count],main=k,col =cbbPalette[7], xlab=expression(bold(r[0][A]/r[0][B])))
count<-count+1
}
mtext("Diversification 0A/0B - no outliers", side =3, line = -1.25, outer = TRUE)
#dev.off()

# for 1A/1B
#png("r_1A_1B.png")
par(mfrow=c(3,3))
count<-1
for (k in k.vec){
hist(div_1A_1B[,count],main=k,col =cbbPalette[8],xlab=expression(bold(r[1][A]/r[1][B])))
count<-count+1
}
mtext("Diversification 1A/1B - no outliers", side =3, line = -1.25, outer = TRUE)
#dev.off()
```
